\clearpage

\renewcommand{\ChapTitle}{\CSPBML}
\renewcommand{\SectionTitle}{CSPB 4622}

\chapter{\ChapTitle}
\section{\SectionTitle}
\horizontalline{0}{0}

\subsection{Course Overview}

The overview of this course can be seen below.

\course{\hypertarget{CSPB:4622}{\CSPBML} \hspace*{1pt} - Prerequisites: \hyperlink{CSPB:2270}{CSPB 2270}, \hyperlink{CSPB:2820}{CSPB 2820}, \hyperlink{CSPB:2824}{CSPB 2824}, \hyperlink{CSPB:3022}{CSPB 3022} - Credits: 3}{
    \vspace*{-1.5em}
    \begin{highlight}[\CSPBML]
        \subsubsection*{Brief Description of Course Content}

        This course provides a comprehensive introduction to machine learning, covering key concepts, methodologies, and applications. Topics include supervised learning, unsupervised learning, neural 
        networks, deep learning, and reinforcement learning. Emphasis is placed on both the theoretical foundations and practical implementations. \vspace*{1em}
        
        \subsubsection*{Specific Outcomes of Instruction}
        
        \begin{itemize}
            \item An ability to explain the fundamental concepts of machine learning and its various applications.
            \item An ability to implement machine learning algorithms in code.
            \item An ability to evaluate the performance of different machine learning models.
            \item An ability to select and apply appropriate machine learning techniques to specific problems.
            \item An ability to understand and utilize neural networks and deep learning architectures.
        \end{itemize}
        
        \subsubsection*{Brief List of Topics to be Covered}
        
        \begin{itemize}
            \item Supervised Learning
            \begin{itemize}
                \item Linear Regression
                \item Logistic Regression
                \item Support Vector Machines
                \item k-Nearest Neighbors
                \item Decision Trees and Random Forests
            \end{itemize}
            \item Unsupervised Learning
            \begin{itemize}
                \item k-Means Clustering
                \item Hierarchical Clustering
                \item Principal Component Analysis (PCA)
            \end{itemize}
            \item Neural Networks
            \begin{itemize}
                \item Perceptrons
                \item Multi-Layer Perceptrons (MLPs)
                \item Backpropagation
            \end{itemize}
            \item Deep Learning
            \begin{itemize}
                \item Convolutional Neural Networks (CNNs)
                \item Recurrent Neural Networks (RNNs)
                \item Applications in computer vision, natural language processing
            \end{itemize}
            \item Reinforcement Learning
            \begin{itemize}
                \item Markov Decision Processes (MDPs)
                \item Q-Learning
                \item Deep Q-Networks (DQNs)
            \end{itemize}
        \end{itemize}
        
        \subsubsection*{Mathematical Concepts Used}
        
        \begin{itemize}
            \item Linear Algebra: Matrices, Vectors, Eigenvalues/Eigenvectors
            \item Probability and Statistics: Probability Distributions, Expectation, Variance
            \item Calculus: Derivatives, Integrals, Gradients, Hessians
            \item Optimization: Gradient Descent, Stochastic Gradient Descent
        \end{itemize}    
    \end{highlight}
}