\clearpage

\renewcommand{\ChapTitle}{\CSPBAlgo}
\renewcommand{\SectionTitle}{CSPB 3104}

\chapter{\ChapTitle}
\section{\SectionTitle}
\horizontalline{0}{0}

\subsection{Course Overview}

The overview of this course can be seen below.

\course{\hypertarget{CSPB:3104}{\CSPBAlgo} \hspace*{1pt} - Prerequisites: \hyperlink{CSPB:2270}{CSPB 2270} \& \hyperlink{CSPB:2824}{CSPB 2824} - Credits: 4}{
    \vspace*{-1.5em}
    \begin{highlight}[\CSPBAlgo]
        \subsubsection*{Brief Description of Course Content}
    
        Covers the fundamentals of algorithms and various algorithmic strategies, including time and space complexity, sorting algorithms, recurrence relations, divide and conquer algorithms, greedy 
        algorithms, dynamic programming, linear programming, graph algorithms, problems in P and NP, and approximation algorithms. \vspace*{1em}
    
        \subsubsection*{Specific Outcomes of Instruction}
    
        \begin{itemize}
            \item Understanding properties of algorithms.
            \item Proving these properties mathematically.
            \item Proving rigorous time and space complexity bounds on the performance.
            \item Understand the relative merits or demerits of an algorithm, in practice.
            \item Use algorithms to solve core problems that may arise inside applications.
            \item Learn key tricks (motifs) underlying the design of new algorithms for emerging applications.
        \end{itemize}
    
        \subsubsection*{Brief List of Topics to be Covered}
    
        \begin{itemize}
            \item Introduction to Algorithms: Complexity analysis
            \item Divide and Conquer Algorithms
            \item Sorting and Order Statistics
            \item Advanced Data Structures: heaps, balanced trees and hash-functions
            \item Dynamic Programming
            \item Greedy Algorithms
            \item Graph Algorithms: Search, Minimum Spanning Trees, Shortest Paths, Network Flows
            \item Introduction to Linear and Integer Programming
            \item Basics of Computational Complexity: P, NP, reductions and open problems
        \end{itemize}
    
        \subsubsection*{Mathematical Concepts Used}
        
        \begin{itemize}
            \item Recursion
            \item Logarithmic
            \item Exponential Functions
            \item Induction
            \item Graph Theory
        \end{itemize}
    \end{highlight}
}

\subsection{Course Notes}

The notes for this course can be found below: \coursedoc{\pdflink{\CSPBAlgoDir CSPB 3104 - Algorithms Course Notes.pdf}{CSPB 3104 - Algorithms Course Notes}}