\clearpage

\renewcommand{\ChapTitle}{\CSPBOS}
\renewcommand{\SectionTitle}{CSPB 3753}

\chapter{\ChapTitle}
\section{\SectionTitle}
\horizontalline{0}{0}

\subsection{Course Overview}

The overview of this course can be seen below.

\course{\hypertarget{CSPB:3753}{\CSPBOS} \hspace*{1pt} - Prerequisites: \hyperlink{CSPB:2270}{CSPB 2270}, \hyperlink{CSPB:2824}{CSPB 2824}, \hyperlink{CSPB:3308}{CSPB 3308} - Credits: 4}{
    \vspace*{-1.5em}
    \begin{highlight}[\CSPBOS]
        \subsubsection*{Brief Description of Course Content}
    
        Examines the structure and function of operating systems as an intermediary between applications and computer hardware. Topics include OS design goals, hardware management, multitasking, process 
        and thread abstractions, file and memory management, security, and networking. Upon completion, students should be able to perform operating systems functions at the support level in a single-user 
        environment. Microcontrollers are ubiquitous in the modern world, in everything from your toaster, microwave, and refrigerator, to the complex and sophisticated systems in satellites and self-driving 
        vehicles.  We have augmented our Operating Systems course to use a Raspberry Pi in hands-on assignments such as adding systems calls to the Linux operating system running on the Raspberry Pi. 
        The goal is for you to learn to apply the theory of operating systems and gain experience physically working with and changing a real working computer. \vspace*{1em}
        
        \subsubsection*{Specific Outcomes of Instruction}
    
        \begin{itemize}
            \item Explain basic concepts in the design and structure of operating systems, including kernel/user mode, system calls, preemptive multitasking, and monolithic/microkernel structure.
            \item Describe how interrupt-based processing achieves efficient management of device input/output communication.
            \item Define processes and threads, describe different ways to communicate between processes and threads, and apply mutual exclusion-based solutions to synchronize multi-threaded processes without deadlock occurring.
            \item Identify different scheduling algorithms and their suitability for different types of applications, including compute-bound, I/O-bound and real time.
            \item Explain the concept of virtual memory, the rationale for on-demand paging, and the role of working sets to avoid thrashing in a caching-based memory hierarchy.
            \item Demonstrate understanding of fundamental concepts in file system design, including linked and indexed file allocation, mounting, a virtual file system layer, memory mapping, journaling, and performance optimizations for storage media (magnetic and solid state).
            \item Describe basic concepts to secure and protect operating systems.
            \item Explain basic concepts in networked operating systems design, including layered network architecture and distributed file systems structure.
            \item Describe the basic concept of a virtual machine and different types of virtual machines.
            \item Successfully modify, add functionality to, and re-compile the kernel of an operating system.
        \end{itemize}
        
        \subsubsection*{Brief List of Topics to be Covered}
    
        \begin{itemize}
            \item Operating System Design and Structure
            \item Device Input/Output Management
            \item Process and Thread Management and Scheduling
            \item Inter-Process Communication and Synchronization
            \item Memory Management
            \item File System and Storage Management
            \item Security
            \item Networking
            \item Virtual Machines
        \end{itemize}
        
        \subsubsection*{Mathematical Concepts Used}
        
        \begin{itemize}
            \item Basic Arithmatic
            \item Averages
            \item Recursions
            \item Random Numbers and Distributions
        \end{itemize}
    \end{highlight}
}