\clearpage

\renewcommand{\ChapTitle}{\CSPBDisc}
\renewcommand{\SectionTitle}{CSPB 2824}

\chapter{\ChapTitle}
\section{\SectionTitle}
\horizontalline{0}{0}

\subsection{Course Overview}

The overview of this course can be seen below.

\course{\hypertarget{CSPB:2824}{\CSPBDisc} \hspace*{1pt} - Co-Requisites: \hyperlink{CSPB:1300}{CSPB 1300} - Credits: 3}{
    \vspace*{-1.5em}
    \begin{highlight}[\CSPBDisc]
        \subsubsection*{Brief Description of Course Content}
    
        The course covers fundamental ideas from discrete mathematics, especially for computer science students. It focuses on topics that will be foundational for future courses including algorithms, 
        artificial intelligence, programming languages, theoretical computer science, computer systems, cryptography, networks, computer/network security, databases, and compilers. \vspace*{1em}
    
        \subsubsection*{Specific Outcomes of Instruction}
    
        We will build on the following 6 primary learning goals throughout the term:
        \begin{itemize}
            \item Understand and construct logical arguments and proofs using formal logic, truth tables, and proof techniques.
            \item Understand and use the basics structures of sets, functions, sums and matrices.
            \item Use and understand algorithms, number theory and cryptography.
            \item Demonstrate and make arguments using counting, and probability.
            \item Use, develop, and analyze formal relations, and graph theory.
            \item Develop the skills of “Mathematical Maturity” including:
            \begin{itemize}
                \item The capacity to generalize from a specific example to broad concept.
                \item The capacity to handle increasingly abstract ideas.
                \item A significant shift from learning by memorization to learning through understanding.
                \item The ability to recognize mathematical patterns and think abstractly.
                \item Read, write and critique formal proofs.
                \item Teach yourself and fill in missing details.
            \end{itemize}
        \end{itemize}
    
        \subsubsection*{Brief List of Topics to be Covered}
    
        \begin{itemize}
            \item Logic
            \item Proof techniques
            \item Algorithms
            \item Modular Arithmetic
            \item Number theory
            \item Cryptography
            \item Induction
            \item Combinatorics
            \item Probability
            \item Bayes Thm
            \item Relations
            \item Graphs
        \end{itemize}
    
        \subsubsection*{Mathematical Concepts Used}
    
        \begin{itemize}
            \item Basic Algebra
            \item Program Entry Requirements
        \end{itemize}
    \end{highlight}
}

\subsection{Course Notes}

The notes for this course can be found below: \coursedoc{\pdflink{\CSPBDiscreteDir CSPB 2824 - Discrete Structures Course Notes.pdf}{CSPB 2824 - Discrete Structures Course Notes}}

\subsection{Course Syllabus}

The syllabus for this course can be found below: \coursedoc{\pdflink{\CSPBDiscreteSyllabusDir CSPB 2824 - Discrete Structures Syllabus.pdf}{CSPB 2824 - Discrete Structures Syllabus}}

\subsection{Course Textbook}

The textbook for this course can be found below: \coursedoc{\pdflink{\CSPBDiscreteDirTextbookDir Discrete Mathematics And Its Applications (7th Edition) - Kenneth Rosen.pdf}{Discrete Mathematics And Its Applications (7th Edition) - Kenneth Rosen}}