\clearpage

\renewcommand{\ChapTitle}{\CSPBDataSci}
\renewcommand{\SectionTitle}{CSPB 3022}

\chapter{\ChapTitle}
\section{\SectionTitle}
\horizontalline{0}{0}

\subsection{Course Overview}

The overview of this course can be seen below.

\course{\hypertarget{CSPB:3022}{\CSPBDataSci} \hspace*{1pt} - Prerequisites: \hyperlink{CSPB:1300}{CSPB 1300} - Credits: 3}{
    \vspace*{-1.5em}
    \begin{highlight}[\CSPBDataSci]
        \subsubsection*{Brief Description of Course Content}
    
        Introduces students to the tools methods and theory behind extracting insights from data. Covers algorithms of cleaning and munging data, probability theory and common distributions, statistical 
        simulation, drawing inferences from data, and basic statistical modeling. \vspace*{1em}
        
        \subsubsection*{Specific Outcomes of Instruction}
    
        \begin{itemize}
            \item Recognize the importance of data collection, identify limitations in data collection methods and other sources of statistical bias, and determine their implications and how they affect the scope of inference.
            \item Use statistical software to summarize data numerically and visually, and to perform data analysis.
            \item Have a conceptual understanding of the unified nature of statistical inference.
            \item Apply estimation and testing methods to analyze single variables or the relationship between two variables in order to understand natural phenomena and make data-based decisions.
            \item Model numerical response variables using a single explanatory variable or multiple explanatory variables in order to investigate relationships between variables.
            \item Interpret results correctly, effectively, and in context without relying on statistical jargon.
            \item Critique data-based claims and evaluate data-based decisions.
        \end{itemize}
        
        \subsubsection*{Brief List of Topics to be Covered}
    
        \begin{itemize}
            \item Data Exploration and Probability
            \item Conditional probability and Bayes rule
            \item Discrete/continuous random variables and computing with distributions
            \item Joint distributions, covariance, correlation and sums of random variables
            \item Using Jupyter python environment
            \item Python tools for data science - NumPy and Pandas
            \item Basic statistical estimation, random samples, bootstrap and resampling techniques, unbiased estimators and confidence intervals for measure data
            \item T-Test
            \item Linear Regression and classification
            \item Maximum likelihood estimation and analysis of variance
        \end{itemize}
        
        \subsubsection*{Mathematical Concepts Used}
        
        \begin{itemize}
            \item Counting Theory
            \item Probabilities
            \item Integration
        \end{itemize}
    \end{highlight}
}

\subsection{Course Notes}

The notes for this course can be found below: \coursedoc{\CSPBDataSciCourseNotes}

\subsection{Course Syllabus}

The syllabus for this course can be found below: \coursedoc{\CSPBDataSciSyllabus}

\subsection{Course Textbook}

The textbooks for this course can be found below: \coursedoc{{\scriptsize{\CSPBDataSciLearnDSTextbook}}} \coursedoc{{\small{\CSPBDataSciPyForDSTextbook}}}