\clearpage

\renewcommand{\ChapTitle}{Week 7: (2/26 - 3/3)}
\renewcommand{\SectionTitle}{Expected Value, Variance - Discrete And Continuous RV}

\chapter{\ChapTitle}

\section{\SectionTitle}
\horizontalline{0}{0}

\subsection{Piazza}

Must post / respond to two Piazza posts. \assignment{3/5/24}{Piazza7DueDate}

\subsection{Lectures}

The lecture videos for this week are:

\begin{itemize}
    \item \lecture{https://applied.cs.colorado.edu/pluginfile.php/76404/mod_resource/content/2/lab6_video.mp4}{Random Variables And Distributions Notebook Walkthrough}{83}
    \item \lecture{https://applied.cs.colorado.edu/pluginfile.php/76405/mod_resource/content/1/CSCI3022_002_CASEE240_10_4_2023.mp4}{RV And Distributions, Expectation And Variance}{54}
    \item \lecture{https://applied.cs.colorado.edu/pluginfile.php/76407/mod_resource/content/1/CSCI3022_002_CASEE240_10_6_2023.mp4}{Joint Distributions, Covariance, Independence}{54}
    \item \lecture{https://applied.cs.colorado.edu/pluginfile.php/76421/mod_resource/content/2/CSCI3022_002_CASEE240_10_9_2023.mp4}{Covariance And Correlation, Independence}{54}
\end{itemize}

\noindent The lecture notes for this week are:

\begin{itemize}
    \item \pdflink{\LecNoteDir Expected Value, Variance Discrete And Continuous RV Lecture Notes.pdf}{Expected Value, Variance Discrete And Continuous RV Lecture Notes}
    \item \pdflink{\LecNoteDir More With RV -  Discrete Vs. Continuous - Expected Value And Variance Lecture Notes.pdf}{More With RV -  Discrete Vs. Continuous - Expected Value And Variance Lecture Notes}
    \item \pdflink{\LecNoteDir Joint Distributions - Covariance And Correlation Lecture Notes.pdf}{Joint Distributions - Covariance And Correlation Lecture Notes}
    \item \pdflink{\LecNoteDir Sampling Lecture Notes.pdf}{Sampling Lecture Notes}
\end{itemize}

\subsection{Assignments}

The assignment for this week is:

\begin{itemize}
    \item The assignment for this week is \textbf{Expected Value, Variance - Discrete And Continuous RV}. \assignment{3/5/24}{Ass6DueDate}
\end{itemize}

\subsection{Concept Summary}

\begin{notes}{Expected Value, Variance - Discrete And Continuous RV}
    \subsection*{Overview}

    Expected Value and Variance are fundamental statistical measures that describe the distribution of random variables (RVs). They play crucial roles in the fields of probability, statistics, and 
    many practical applications like finance and risk management. Expected value provides a measure of the central tendency of a random variable, while variance measures the spread or dispersion 
    around this central value. Understanding these concepts for both discrete and continuous random variables allows for deeper insights into the behavior of random processes and decision-making 
    under uncertainty. \vspace*{1em}
    
    \subsubsection*{Expected Value}
    
    The Expected Value (EV) of a random variable gives a sense of the 'average' outcome one might expect if an experiment were repeated many times:
    \begin{itemize}
        \item \textbf{Discrete Random Variables}: The expected value is calculated by summing the products of each possible value of the random variable and its corresponding probability. It is denoted 
        as $E(X) = \sum_{i=1}^n x_i p_i$, where $x_i$ are the possible values and $p_i$ their probabilities.
        \item \textbf{Continuous Random Variables}: For continuous variables, the expected value is determined by integrating the product of the variable's values and their probability density function 
        (PDF) over the possible range of values. It is expressed as $E(X) = \int_{-\infty}^{\infty} x f(x) \, dx$, where $f(x)$ is the PDF.
    \end{itemize}
    
    \subsubsection*{Variance}
    
    Variance quantifies the variability or spread of a random variable's possible outcomes around its expected value:
    \begin{itemize}
        \item \textbf{Discrete Random Variables}: Variance is calculated by summing the squared differences between each possible value and the expected value, each weighted by its probability: $\text{Var}(X) = \sum_{i=1}^n (x_i - E(X))^2 p_i$.
        \item \textbf{Continuous Random Variables}: Similar to the discrete case, but involving integration: $\text{Var}(X) = \int_{-\infty}^{\infty} (x - E(X))^2 f(x) \, dx$.
    \end{itemize}
    
    Both expected value and variance are essential for understanding the distribution characteristics of random variables, influencing everything from risk assessment to optimization in various applications.
    
    \subsubsection*{Summary}
    
    Expected Value and Variance are critical in both theoretical and applied statistics, providing key insights into the likelihood and variability of outcomes for discrete and continuous random variables. 
    These measures help statisticians and data scientists to predict outcomes and make informed decisions based on the underlying probabilities. Understanding these concepts is fundamental for analyzing 
    risk, optimizing strategies, and modeling in uncertain environments.    
\end{notes}