\clearpage
\chapter{Week 17: (12/18 - 12/24)}

\section{Final Exam}
\horizontalline{0}{0}

\subsection{Lectures}

The lectures for this week and their links can be found below:

\begin{itemize}
    \item \href{https://applied.cs.colorado.edu/mod/resource/view.php?id=51884}{Final Exam Notes} $\approx$ 2 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51907}{Introduction} $\approx$ 16 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51908}{Introduction to Lists} $\approx$ 5 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51909}{List Operations} $\approx$ 8 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51910}{List Manipulations: Iterations} $\approx$ 8 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51911}{Python Tutor: Getting Started} $\approx$ 5 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51912}{Basics: Defining Functions } $\approx$ 10 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51913}{Defining Functions : Multiple Arguments And Larger Functions} $\approx$ 13 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51914}{Composition of Functions} $\approx$ 8 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51915}{Recursive Functions} $\approx$ 10 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51916}{Side Effects: Pure vs. Nonpure Functions} $\approx$ 7 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51917}{Polymorphic Functions} $\approx$ 12 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51918}{Higher Order Functions} $\approx$ 10 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51919}{Lambdas} $\approx$ 7 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51920}{Closures} $\approx$ 18 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51921}{Map Operation} $\approx$ 17 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51922}{Reduce Operation} $\approx$ 12 min.
    \item \href{https://applied.cs.colorado.edu/mod/hvp/view.php?id=51923}{Filter} $\approx$ 10 min.
\end{itemize}

\noindent Below is a list of lecture notes for this week:

\begin{itemize}
    \item \pdflink{./Notes/Introduction To Functional Programming Lecture Notes.pdf}{Introduction To Functional Programming Lecture Notes}
\end{itemize}

\noindent The Python tutor links for this final assignment can be found below:

\begin{itemize}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Lists}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{List Operations}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Iterations}
    \item \href{https://pythontutor.com/visualize.html#mode=display}{Introduction}
    \item \href{https://pythontutor.com/visualize.html#mode=display}{Defining Functions}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Multiple Argument Functions}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Composition Of Functions}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Recursive Functions}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Pure Functions}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Nonpure Functions}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Polymorphic Functions}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Higher Order Functions}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Lambda Functions}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Closures}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Maps: 1}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Maps: 2}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Maps: 3}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Reduce Operations: 1}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Reduce Operations: 2}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Filters: 1}
    \item \href{https://pythontutor.com/visualize.html#mode=edit}{Filters: 2}
\end{itemize}

\subsection{Exam}

The exam for this week can be found at \href{https://applied.cs.colorado.edu/mod/quiz/view.php?id=52201}{Final Exam} \due{(12/18/23)} \cbox{exam-week17} \vspace*{1em}

The optional final programming assignment this week can be found at \href{https://applied.cs.colorado.edu/mod/quiz/view.php?id=51924}{Final Programming Assignment} \due{(12/16/23)} \cbox{(proj-week17)}